### 2025-06-24 sections/2-methods.tex
formatting fixes

\begin{algorithm} % enter the algorithm environment
    \caption{Partition($C$, $criteria$)} % give the algorithm a caption
    \label{alg:methods:partition} % and a label for \ref{} commands later in the document
    \begin{algorithmic} % enter the algorithmic environment
        \Require $f: X \times X \mapsto \mathbb{R}^+ \cup \{0\}$, a distance function
        \Require $C$, a cluster
        \Require $criteria$, user-specified continuation criteria

        \State $seeds \Leftarrow$ random sample of $\left\lceil \sqrt{|C|} \right\rceil$ points from $C$
        \State $c \Leftarrow$ geometric median of $seeds$
        \State $l \Leftarrow \argmax f(c, x) \ \forall \ x \in C$
        \State $r \Leftarrow \argmax f(l, x) \ \forall \ x \in C$
        \State $L \Leftarrow \{x \ | \ x \in C \land f(l, x) \le f(r, x)\}$
        \State $R \Leftarrow \{x \ | \ x \in C \land f(r, x) < f(l, x)\}$

        \If{$|L| > 1$ \textbf{and} $L$ satisfies $criteria$}
            \State Partition($L$, $criteria$)
        \EndIf

        \If{$|R| > 1$ \textbf{and} $R$ satisfies $criteria$}
            \State Partition($R$, $criteria$)
        \EndIf
    \end{algorithmic}
\end{algorithm}

@@

\begin{figure}[H]
    \vskip -0.2in
    \centering
    \includegraphics[scale=0.65]{images/geometry/deltas.pdf}
    \caption{
        {\color{blue}$\delta$}, {\color{red}$\delta^{+}$}, and {\color{green}$\delta^{-}$} for a cluster $C$ and a query $q$.
        ${\color{blue}\delta} = f(q, c)$ is the distance from the query to the cluster center $c$.
        ${\color{red}\delta^{+}} = \delta + r$ is the distance from the query to the theoretically farthest point in $C$.
        ${\color{green}\delta^{-}} = \text{max}(0, \delta - r)$ is the distance from the query to the theoretically closest point in $C$.
    }
    \label{fig:methods:deltas}
    \vskip -0.5in
\end{figure}

\begin{minipage}{.425\textwidth}
    \begin{algorithm}[H]
        \caption{tree-search($C$, $q$, $r$)}
        \label{alg:methods:rnn-search:tree-search}
        \begin{algorithmic}
            \Require $f$, a distance function
            \Require $C$, a cluster
            \Require $q$, a query
            \Require $r$, a search radius
            \If{$\delta^+_C \leq r$}
                \State \textbf{return} $\{C\}$
            \Else
                \State $[L, R]$ $\Leftarrow$ \textit{children} of $C$
                \State \textbf{return} tree-search($L, q, r$) \\
                \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ $\cup$ tree-search($R, q, r$)
            \EndIf
        \end{algorithmic}
    \end{algorithm}
\end{minipage}
\hfill
\begin{minipage}{.475\textwidth}
    \begin{algorithm}[H]
        \caption{leaf-search($Q$, $q$, $r$)}
        \label{alg:methods:rnn-search:leaf-search}
        \begin{algorithmic}
            \Require $f$, a distance function
            \Require $Q$, a set of clusters
            \Require $q$, a query
            \Require $r$, a search radius
            \State $H \Leftarrow \emptyset$
            \For{$C \in Q$}
                \If{$\delta^+_C \leq r$}
                    \State $H$ $\Leftarrow$ $H \cup C$
                \Else
                    \For{$p \in C$}
                        \If{$f(p, q) \leq r$}
                            \State $H$ $\Leftarrow$ $H \cup \{p\}$
                        \EndIf
                    \EndFor
                \EndIf
            \EndFor
            \State \textbf{return} $H$
        \end{algorithmic}
    \end{algorithm}
\end{minipage}

@@

\begin{algorithm} % enter the algorithm environment
    \caption{Repeated $\rho$-NN($\mathcal{R}$, $q$, $k$)} % give the algorithm a caption
    \label{alg:methods:repeated-rnn} % and a label for \ref{} commands later in the document
    \begin{algorithmic} % enter the algorithmic environment
        \Require $\mathcal{R}$, the root cluster
        \Require $q$, a query
        \Require $k$, the number of neighbors to find
        \State{$H \Leftarrow [\ ]$, a max-heap by $\delta$ of size $k$}
        \State $r \Leftarrow radius$ of $\mathcal{R}$
        \State $r \Leftarrow$ $\frac{r}{|\mathcal{R}|}$
        \State $Q \Leftarrow$ tree-search($\mathcal{R}$, $q$, $r$)
        \While{$\sum_{C \in Q} |C| < k$}
            \If{$Q = \emptyset$}
                \State $r \Leftarrow 2 \cdot r$
            \ElsIf{$\sum_{C \in Q} |C| >= k$}
                \State $\mu \Leftarrow \frac{1}{|Q|} \cdot \sum_{C \in Q} \big( LFD(C)^{-1} \big)$
                \State $r \Leftarrow r \cdot \min \bigg( 2, \left( {\frac{k}{\sum_{C \in Q} |C|}} \right)^{\mu} \bigg)$
            \EndIf
            \State $Q \Leftarrow$ tree-search($\mathcal{R}$, $q$, $r$)
        \EndWhile
        \State $H \Leftarrow \text{leaf-search}(Q, q, r)$
        \State \textbf{return} $H$ as a list
    \end{algorithmic}
\end{algorithm}

@@

\begin{theorem} Let $X$ be a dataset and $q$ a query sampled from the same distribution (i.e., arising from the same generative process) as $X$. Then time complexity of performing Repeated $\rho$-NN search on $X$ with query $q$ is \begin{gather}
        \mathcal{O}
        \Bigg(
            \underbrace{
                \log~\overbrace{\mathcal{N}_{\hat{r}}(X)}^{\textrm{metric entropy}}
            }_{\textrm{tree-search}}
            \ + \
            \underbrace{
                \overbrace{k}^{\textrm{output size}} \cdot
                \overbrace{\bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d}^{\textrm{scaling factor}}
            }_{\textrm{leaf-search}}
        \Bigg)
        \label{eq:methods:repeated-rnn-complexity}
    \end{gather}
    where $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy of the dataset, $d$ is the LFD of the dataset, and $k$ is the number of nearest neighbors.
    \label{thm:methods:rnn-complexity}
\end{theorem}

@@

\begin{minipage}{0.5125\textwidth}
    \begin{algorithm}[H]\small
        \caption{Breadth-First Sieve($\mathcal{R}$, $q$, $k$)} % give the algorithm a caption
        \label{alg:methods:bredth-first-sieve} % and a label for \ref{} commands later in the document
        \begin{algorithmic} % enter the algorithmic environment
            \Require $\mathcal{R}$, the root cluster
            \Require $q$, a query
            \Require $k$, the number of neighbors to find
            \State $c \Leftarrow$ \textit{center} of $\mathcal{R}$
            \State $Q \Leftarrow$ \{ ($\mathcal{R}$, $\delta^{+}_{\mathcal{R}}$, $|\mathcal{R}| - 1$), ($c$, $\delta_{\mathcal{R}}$, 1) \}
            \While{$\sum_{(\_, \_, m) \in Q} m \neq k$}
                \State $\tau \Leftarrow$ QuickSelect($Q$, $k$)
                \State $Q^{'} \Leftarrow \emptyset$
                \For{$(C, \_, \_) \in Q$}
                    \If{$\delta^{-}_{C} \leq \tau$}
                        \If{$C$ is a point}
                            \State $Q^{'} \Leftarrow Q^{'} \cup \{ (C, \delta_{C}, 1) \}$
                        \ElsIf{$C$ is a leaf}
                            \State $Q^{'} \Leftarrow Q^{'} \cup \{ (p, \delta_{p}, 1)$ for $p \in C \}$
                        \Else
                            \State $[L, R] \Leftarrow$ children of $C$
                            \State $l, r \Leftarrow$ centers of $L, R$
                            \State $Q^{'} \Leftarrow Q^{'} \cup \{ (L, \delta^{+}_{L}, |L| - 1), (l, \delta_{L}, 1) \}$
                            \State $Q^{'} \Leftarrow Q^{'} \cup \{ (R, \delta^{+}_{R}, |R| - 1), (l, \delta_{R}, 1) \}$
                        \EndIf
                    \EndIf
                \EndFor
                \State $Q \Leftarrow Q^{'}$
            \EndWhile
            \State QuickSelect($Q$, $k$)
            \State \textbf{return} The first $k$ points in $Q$
        \end{algorithmic}
    \end{algorithm}
\end{minipage}
\hfill
\begin{minipage}{0.425\textwidth}
    \begin{algorithm}[H]\small
        \caption{Depth-First Sieve($\mathcal{R}$, $q$, $k$)}
        \label{alg:methods:depth-first-sieve}
        \begin{algorithmic}
            \Require $\mathcal{R}$, the root cluster
            \Require $q$, a query
            \Require $k$, the number of neighbors to find
            \State{$Q \Leftarrow [\mathcal{R}]$, a min-heap by $\delta^{-}$}
            \State{$H \Leftarrow [\ ]$, a max-heap by $\delta$ of size $k$}
            \While{$|H| < k$ \textbf{or} $H.peek.\delta \geq Q.peek.\delta^{-}$}
                \While{$Q.peek$ is not a leaf}
                    \State{$C \Leftarrow Q.pop$, the closest cluster}
                    \State{$[L, R] \Leftarrow$ children of $C$}
                    \State{$Q.push(L)$}
                    \State{$Q.push(R)$}
                \EndWhile
                \State{$leaf \Leftarrow Q.pop$}
                \For{$p \in leaf$}
                    \State{$H.push(p)$}
                \EndFor
            \EndWhile
            \State \textbf{return} $H$
        \end{algorithmic}
    \end{algorithm}
\end{minipage}

@@

\begin{theorem}
    Let $X$ be a dataset and $q$ a query sampled from the same distribution (i.e., arising from the same generative process) as $X$. Let $T \coloneqq \mathcal{O} \big( \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(X) \big)$ and   $L \coloneqq \mathcal{O} \left( k \cdot \bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d \right)$, where $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy of $X$, $d$ is the LFD of $X$, $\hat{|C|}$ is the mean cardinality of clusters overlapping the query ball, and $k$ is the number of nearest neighbors. Then, for dataset $X$ and query $q$, the time complexity of performing Breadth-First Sieve search is \begin{equation}
        \mathcal{O} \Big( (T + L ) \log (T + L ) \Big)
        \label{eq:methods:breadth-first-sieve-complexity}
    \end{equation} and the the time complexity of performing Depth-First Sieve search is \begin{equation}
        \mathcal{O} \Big( T \log T + L \log k \Big).
        \label{eq:methods:depth-first-sieve-complexity}
    \end{equation}
    \label{thm:methods:sieve-complexity}
\end{theorem}

>>>>

\begin{algorithm} % enter the algorithm environment
    \caption{Partition($C$, $criteria$)} % give the algorithm a caption
    \label{alg:methods:partition} % and a label for \ref{} commands later in the document
    \begin{algorithmic} % enter the algorithmic environment
    \REQUIRE $f: X \times X \mapsto \mathbb{R}^+ \cup \{0\}$, a distance function
    \REQUIRE $C$, a cluster
    \REQUIRE $criteria$, user-specified continuation criteria

    \STATE $seeds \Leftarrow$ random sample of $\left\lceil \sqrt{|C|} \right\rceil$ points from $C$
    \STATE $c \Leftarrow$ geometric median of $seeds$
    \STATE $l \Leftarrow \argmax f(c, x) \ \forall \ x \in C$
    \STATE $r \Leftarrow \argmax f(l, x) \ \forall \ x \in C$
    \STATE $L \Leftarrow \{x \ | \ x \in C \land f(l, x) \le f(r, x)\}$
    \STATE $R \Leftarrow \{x \ | \ x \in C \land f(r, x) < f(l, x)\}$

    \IF{$|L| > 1$ \textbf{and} $L$ satisfies $criteria$}
        \STATE Partition($L$, $criteria$)
    \ENDIF

    \IF{$|R| > 1$ \textbf{and} $R$ satisfies $criteria$}
        \STATE Partition($R$, $criteria$)
    \ENDIF
\end{algorithmic}
\end{algorithm}

@@

% \begin{figure}[H]
%     \vskip -0.2in
%     \centering
%     \includegraphics[scale=0.65]{images/geometry/deltas.pdf}
%     \caption{
%         {\color{blue}$\delta$}, {\color{red}$\delta^{+}$}, and {\color{green}$\delta^{-}$} for a cluster $C$ and a query $q$.
%         ${\color{blue}\delta} = f(q, c)$ is the distance from the query to the cluster center $c$.
%         ${\color{red}\delta^{+}} = \delta + r$ is the distance from the query to the theoretically farthest point in $C$.
%         ${\color{green}\delta^{-}} = \text{max}(0, \delta - r)$ is the distance from the query to the theoretically closest point in $C$.
%     }
%     \label{fig:methods:deltas}
%     \vskip -0.5in
% \end{figure}

\begin{minipage}{.425\textwidth}
    \begin{algorithm}[H]
        \caption{tree-search($C$, $q$, $r$)}
        \label{alg:methods:rnn-search:tree-search}
        \begin{algorithmic}
            \REQUIRE $f$, a distance function
            \REQUIRE $C$, a cluster
            \REQUIRE $q$, a query
            \REQUIRE $r$, a search radius
            \IF{$\delta^+_C \leq r$}
                \STATE \textbf{return} $\{C\}$
            \ELSE
                \STATE $[L, R]$ $\Leftarrow$ \textit{children} of $C$
                \STATE \textbf{return} tree-search($L, q, r$) \\
                \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ $\cup$ tree-search($R, q, r$)
            \ENDIF
        \end{algorithmic}
    \end{algorithm}
\end{minipage}
\hfill
\begin{minipage}{.475\textwidth}
    \begin{algorithm}[H]
        \caption{leaf-search($Q$, $q$, $r$)}
        \label{alg:methods:rnn-search:leaf-search}
        \begin{algorithmic}
            \REQUIRE $f$, a distance function
            \REQUIRE $Q$, a set of clusters
            \REQUIRE $q$, a query
            \REQUIRE $r$, a search radius
            \STATE $H \Leftarrow \emptyset$
            \FOR{$C \in Q$}
                \IF{$\delta^+_C \leq r$}
                    \STATE $H$ $\Leftarrow$ $H \cup C$
                \ELSE
                    \FOR{$p \in C$}
                        \IF{$f(p, q) \leq r$}
                            \STATE $H$ $\Leftarrow$ $H \cup \{p\}$
                        \ENDIF
                    \ENDFOR
                \ENDIF
            \ENDFOR
            \STATE \textbf{return} $H$
        \end{algorithmic}
    \end{algorithm}
\end{minipage}

@@

\begin{algorithm} % enter the algorithm environment
    \caption{Repeated $\rho$-NN($\mathcal{R}$, $q$, $k$)} % give the algorithm a caption
    \label{alg:methods:repeated-rnn} % and a label for \ref{} commands later in the document
    \begin{algorithmic} % enter the algorithmic environment
        \REQUIRE $\mathcal{R}$, the root cluster
        \REQUIRE $q$, a query
        \REQUIRE $k$, the number of neighbors to find
        \STATE{$H \Leftarrow [\ ]$, a max-heap by $\delta$ of size $k$}
        \STATE $r \Leftarrow radius$ of $\mathcal{R}$
        \STATE $r \Leftarrow$ $\frac{r}{|\mathcal{R}|}$
        \STATE $Q \Leftarrow$ tree-search($\mathcal{R}$, $q$, $r$)
        \WHILE{$\sum_{C \in Q} |C| < k$}
            \IF{$Q = \emptyset$}
                \STATE $r \Leftarrow 2 \cdot r$
            \ELSIF{$\sum_{C \in Q} |C| >= k$}
                \STATE $\mu \Leftarrow \frac{1}{|Q|} \cdot \sum_{C \in Q} \big( LFD(C)^{-1} \big)$
                \STATE $r \Leftarrow r \cdot \min \bigg( 2, \left( {\frac{k}{\sum_{C \in Q} |C|}} \right)^{\mu} \bigg)$
            \ENDIF
            \STATE $Q \Leftarrow$ tree-search($\mathcal{R}$, $q$, $r$)
        \ENDWHILE
        \STATE $H \Leftarrow \text{leaf-search}(Q, q, r)$
        \STATE \textbf{return} $H$ as a list
    \end{algorithmic}
\end{algorithm}

@@

% \begin{theorem} Let $X$ be a dataset and $q$ a query sampled from the same distribution (i.e., arising from the same generative process) as $X$. Then time complexity of performing Repeated $\rho$-NN search on $X$ with query $q$ is \begin{gather}
%         \mathcal{O}
%         \Bigg(
%             \underbrace{
%                 \log~\overbrace{\mathcal{N}_{\hat{r}}(X)}^{\textrm{metric entropy}}
%             }_{\textrm{tree-search}}
%             \ + \
%             \underbrace{
%                 \overbrace{k}^{\textrm{output size}} \cdot
%                 \overbrace{\bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d}^{\textrm{scaling factor}}
%             }_{\textrm{leaf-search}}
%         \Bigg)
%         \label{eq:methods:repeated-rnn-complexity}
%     \end{gather}
%     where $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy of the dataset, $d$ is the LFD of the dataset, and $k$ is the number of nearest neighbors.
%     \label{thm:methods:rnn-complexity}
% \end{theorem}

@@

    \begin{algorithm}
        \caption{Breadth-First Sieve($\mathcal{R}$, $q$, $k$)} % give the algorithm a caption
        \label{alg:methods:bredth-first-sieve} % and a label for \ref{} commands later in the document
        \begin{algorithmic} % enter the algorithmic environment
            \REQUIRE $\mathcal{R}$, the root cluster
            \REQUIRE $q$, a query
            \REQUIRE $k$, the number of neighbors to find
            \STATE $c \Leftarrow$ \textit{center} of $\mathcal{R}$
            \STATE $Q \Leftarrow$ \{ ($\mathcal{R}$, $\delta^{+}_{\mathcal{R}}$, $|\mathcal{R}| - 1$), ($c$, $\delta_{\mathcal{R}}$, 1) \}
            \WHILE{$\sum_{(\_, \_, m) \in Q} m \neq k$}
                \STATE $\tau \Leftarrow$ QuickSelect($Q$, $k$)
                \STATE $Q^{'} \Leftarrow \emptyset$
                \FOR{$(C, \_, \_) \in Q$}
                    \IF{$\delta^{-}_{C} \leq \tau$}
                        \IF{$C$ is a point}
                            \STATE $Q^{'} \Leftarrow Q^{'} \cup \{ (C, \delta_{C}, 1) \}$
                        \ELSIF{$C$ is a leaf}
                            \STATE $Q^{'} \Leftarrow Q^{'} \cup \{ (p, \delta_{p}, 1)$ for $p \in C \}$
                        \ELSE
                            \STATE $[L, R] \Leftarrow$ children of $C$
                            \STATE $l, r \Leftarrow$ centers of $L, R$
                            \STATE $Q^{'} \Leftarrow Q^{'} \cup \{ (L, \delta^{+}_{L}, |L| - 1), (l, \delta_{L}, 1) \}$
                            \STATE $Q^{'} \Leftarrow Q^{'} \cup \{ (R, \delta^{+}_{R}, |R| - 1), (l, \delta_{R}, 1) \}$
                        \ENDIF
                    \ENDIF
                \ENDFOR
                \STATE $Q \Leftarrow Q^{'}$
            \ENDWHILE
            \STATE QuickSelect($Q$, $k$)
            \STATE \textbf{return} The first $k$ points in $Q$
        \end{algorithmic}
    \end{algorithm}

    \begin{algorithm}
        \caption{Depth-First Sieve($\mathcal{R}$, $q$, $k$)}
        \label{alg:methods:depth-first-sieve}
        \begin{algorithmic}
            \REQUIRE $\mathcal{R}$, the root cluster
            \REQUIRE $q$, a query
            \REQUIRE $k$, the number of neighbors to find
            \STATE{$Q \Leftarrow [\mathcal{R}]$, a min-heap by $\delta^{-}$}
            \STATE{$H \Leftarrow [\ ]$, a max-heap by $\delta$ of size $k$}
            \WHILE{$|H| < k$ \textbf{or} $H.peek.\delta \geq Q.peek.\delta^{-}$}
                \WHILE{$Q.peek$ is not a leaf}
                    \STATE{$C \Leftarrow Q.pop$, the closest cluster}
                    \STATE{$[L, R] \Leftarrow$ children of $C$}
                    \STATE{$Q.push(L)$}
                    \STATE{$Q.push(R)$}
                \ENDWHILE
                \STATE{$leaf \Leftarrow Q.pop$}
                \FOR{$p \in leaf$}
                    \STATE{$H.push(p)$}
                \ENDFOR
            \ENDWHILE
            \STATE \textbf{return} $H$
        \end{algorithmic}
    \end{algorithm}

@@

% \begin{theorem}
%     Let $X$ be a dataset and $q$ a query sampled from the same distribution (i.e., arising from the same generative process) as $X$. Let $T \coloneqq \mathcal{O} \big( \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(X) \big)$ and   $L \coloneqq \mathcal{O} \left( k \cdot \bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d \right)$, where $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy of $X$, $d$ is the LFD of $X$, $\hat{|C|}$ is the mean cardinality of clusters overlapping the query ball, and $k$ is the number of nearest neighbors. Then, for dataset $X$ and query $q$, the time complexity of performing Breadth-First Sieve search is \begin{equation}
%         \mathcal{O} \Big( (T + L ) \log (T + L ) \Big)
%         \label{eq:methods:breadth-first-sieve-complexity}
%     \end{equation} and the the time complexity of performing Depth-First Sieve search is \begin{equation}
%         \mathcal{O} \Big( T \log T + L \log k \Big).
%         \label{eq:methods:depth-first-sieve-complexity}
%     \end{equation}
%     \label{thm:methods:sieve-complexity}
% \end{theorem}
